---
title: "PDXDataSciStkMkt"
author: "Charles Howard"
date: "October 21, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The raw data

I read the data in using read.csv with defaults:
````{r, include=TRUE}
datadir<-"C:/Users/Charles/Documents/StockStuff/PDXDataSci/"
dat<-read.csv(paste(datadir,"stocks-us-adjClose.csv",sep=""))
# get column wise count of NA's
nacnt<-sapply(1:dim(dat)[2],function(n){length(which(is.na(dat[,n])))})
# looks like I just need to change the name of col 1 to date
names(dat)[1]<-"Date"
dat$Date<-as.Date(as.character(dat$Date))
```

The data provided to the group consists of 12,032 rows by 711 columns.  
  
```{r,include=TRUE}
dim(dat)
```
The first column is a date and the remaining 710 columns are adjusted closing prices for securities.  The date range is:
```{r include=TRUE}
range(dat$Date)
```
Due to the ~ 47 year date range, there is a large variance in number of empty cells column-to-column.  
```{r}
barplot(nacnt,main="Empty Cell Count by Column",las=1,xlab="Column")
```
## Identifying the Securities

Next I gathered some basic information about each of the securities (Company name, Market Sector, and Industry) from Yahoo Finance:

```{r include=TRUE,message=FALSE,warning=FALSE,error=FALSE,cache=TRUE}
library(rvest)
# lookup each symbol and retrieve the sector and industry
# two problems scraping Yahoo Finance.
# 1.) intermittent HTTP error 503
# 2.) successful read_html, but no data on the company
sybls<-names(dat)[2:length(names(dat))]
compdata<-c()
for(i in sybls){
  errflag<-0
  Sys.sleep(0.001)
  e<-simpleError("HTTP error 503.")
  url<-paste("https://finance.yahoo.com/quote/",i,"/profile?p=",i,sep="")
  test<-try(webpg<-read_html(url))
  if(class(test) %in% "try-error"){
    cmpy_name<-NA
    stk_sector<-NA
    stk_industry<-NA
    compdata<-rbind(compdata,cbind(i,cmpy_name,stk_sector,stk_industry))
    next} else
#  webpg<-read_html(url)
  cmpy_name<-html_text(html_nodes(webpg,"div h3[class='Mb(10px)']"))
  strspans<-html_nodes(webpg,"strong , span") 
  strspans_text<-html_text(strspans)
  stk_sector<-strspans_text[which(strspans_text %in% "Sector")+1]
  stk_industry<-strspans_text[which(strspans_text %in% "Industry")+1]
  if(length(cmpy_name)==0){cmpy_name<-NA}
  if(length(stk_sector)==0){stk_sector<-NA}
  if(length(stk_industry)==0){stk_industry<-NA}
  compdata<-rbind(compdata,cbind(i,cmpy_name,stk_sector,stk_industry))
}
# there are 111 sybls have 0 length character objects in
# stk_sector and stk_industry
compdata[which(nchar(compdata[,3])==0),3]<-NA
compdata[which(nchar(compdata[,4])==0),4]<-NA
compdata_df<-as.data.frame(compdata)
names(compdata_df)<-c("Symbol","Company.Name","Sector","Industry")
```
Breakdown of Sectors:
```{r}
oldpar<-par()
nwmar<-par("mar")+c(1,0,0,0)
par(mar=nwmar)
sectortb<-tapply(compdata_df$Sector,compdata_df$Sector,length)
sectortb<-sectortb[order(sectortb,decreasing = TRUE)]
barplot(sectortb,main="Sector Breakdown",las=2,cex.names = 0.7)

```
```{r include=FALSE}
par<-oldpar
```

Sector Analysis

Created a list of time series for each sector and plotted various quantities.

```{r include=TRUE, message=FALSE,cache=TRUE}
library(xts)
sector_rowsums_ls<-lapply(levels(compdata_df$Sector),function(n){
  symb<-compdata_df$Symbol[which(compdata_df$Sector %in% n)]
  cols<-which(names(dat) %in% symb)
  rsm<-unlist(sapply(1:dim(dat)[1],function(n){sum(dat[n,cols],na.rm=T)}))
  rsm_ts<-xts(rsm,order.by = dat$Date)
  zeroes<-which(rsm_ts==0)
  if(length(zeroes)==0){rsm_ts} else 
    {rsm_ts<-rsm_ts[-zeroes]}
})
```
Plots of daily row sums for each sector. Note: Rows containing zeroes have been removed.

```{r,include=TRUE}
  plot(sector_rowsums_ls[[1]],main=paste(levels(compdata_df$Sector)[1],
                                         "\nAdj. Closing Rowsums",sep=""))
  plot(sector_rowsums_ls[[2]],main=paste(levels(compdata_df$Sector)[2],
                                         "\nAdj. Closing Rowsums",sep=""))
  plot(sector_rowsums_ls[[3]],main=paste(levels(compdata_df$Sector)[3],
                                         "\nAdj. Closing Rowsums",sep=""))
  plot(sector_rowsums_ls[[4]],main=paste(levels(compdata_df$Sector)[4],
                                         "\nAdj. Closing Rowsums",sep=""))
  plot(sector_rowsums_ls[[5]],main=paste(levels(compdata_df$Sector)[5],
                                         "\nAdj. Closing Rowsums",sep=""))
  plot(sector_rowsums_ls[[6]],main=paste(levels(compdata_df$Sector)[6],
                                         "\nAdj. Closing Rowsums",sep=""))
  plot(sector_rowsums_ls[[7]],main=paste(levels(compdata_df$Sector)[7],
                                         "\nAdj. Closing Rowsums",sep=""))
  plot(sector_rowsums_ls[[8]],main=paste(levels(compdata_df$Sector)[8],
                                         "\nAdj. Closing Rowsums",sep=""))
  plot(sector_rowsums_ls[[9]],main=paste(levels(compdata_df$Sector)[9],
                                         "\nAdj. Closing Rowsums",sep=""))

```

Plots of daily log-differences by sector.
```{r,include=TRUE}
plot(diff.xts(log(sector_rowsums_ls[[1]])),main=paste(levels(compdata_df$Sector)[1],
                                       "\nAdj. Daily log(Rowsums) Differences",sep=""))
plot(diff.xts(log(sector_rowsums_ls[[2]])),main=paste(levels(compdata_df$Sector)[2],
                                                      "\nAdj. Daily log(Rowsums) Differences",sep=""))
plot(diff.xts(log(sector_rowsums_ls[[3]])),main=paste(levels(compdata_df$Sector)[3],
                                                      "\nAdj. Daily log(Rowsums) Differences",sep=""))
plot(diff.xts(log(sector_rowsums_ls[[4]])),main=paste(levels(compdata_df$Sector)[4],
                                                      "\nAdj. Daily log(Rowsums) Differences",sep=""))
plot(diff.xts(log(sector_rowsums_ls[[5]])),main=paste(levels(compdata_df$Sector)[5],
                                                      "\nAdj. Daily log(Rowsums) Differences",sep=""))
plot(diff.xts(log(sector_rowsums_ls[[6]])),main=paste(levels(compdata_df$Sector)[6],
                                                      "\nAdj. Daily log(Rowsums) Differences",sep=""))
plot(diff.xts(log(sector_rowsums_ls[[7]])),main=paste(levels(compdata_df$Sector)[7],
                                                      "\nAdj. Daily log(Rowsums) Differences",sep=""))
plot(diff.xts(log(sector_rowsums_ls[[8]])),main=paste(levels(compdata_df$Sector)[8],
                                                      "\nAdj. Daily log(Rowsums) Differences",sep=""))
plot(diff.xts(log(sector_rowsums_ls[[9]])),main=paste(levels(compdata_df$Sector)[9],
                                                      "\nAdj. Daily log(Rowsums) Differences",sep=""))

```

